\documentclass[12pt,reqno]{amsart}
\usepackage[top=2cm, left=2cm,right=2cm,bottom=2cm]{geometry}
\renewcommand{\baselinestretch}{1.2}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xspace}
\usepackage{scalefnt}
\usepackage{tikz}
\usepackage{color,hyperref,enumerate,multicol}
\definecolor{darkblue}{rgb}{0.0,0.0,0.3}
\hypersetup{colorlinks,breaklinks,
            linkcolor=darkblue,urlcolor=darkblue,
            anchorcolor=darkblue,citecolor=darkblue}
            
\usepackage{algorithm}
\usepackage{algorithmic}
\pagestyle{empty}
\newcommand{\N}{\ensuremath{\mathbb{N}}}
\newcommand{\Z}{\ensuremath{\mathbb{Z}}}
\newcommand{\R}{\ensuremath{\mathbb{R}}}
\newcommand{\bL}{\ensuremath{\mathbf{L}}}
\newcommand{\bP}{\ensuremath{\mathbf{P}}}
\newcommand{\bQ}{\ensuremath{\mathbf{Q}}}
\newcommand{\bA}{\ensuremath{\mathbf{A}}}
\newcommand{\bB}{\ensuremath{\mathbf{B}}}
\newcommand{\bG}{\ensuremath{\mathbf{G}}}
\newcommand{\bH}{\ensuremath{\mathbf{H}}}
\newcommand{\invG}{\ensuremath{\operatorname{inv}^{\bG}}}
\newcommand{\invH}{\ensuremath{\operatorname{inv}^{\bH}}}
\newcommand{\aut}{\ensuremath{\operatorname{Aut}}}
\newcommand{\inn}{\ensuremath{\operatorname{Inn}}}
\newcommand{\meet}{\ensuremath{\wedge}}
\newcommand{\Meet}{\ensuremath{\bigwedge}}
\newcommand{\<}{\ensuremath{\langle}}
\renewcommand{\>}{\ensuremath{\rangle}}
\newcommand{\join}{\ensuremath{\vee}}
\renewcommand{\emptyset}{\ensuremath{\varnothing}}
\renewcommand{\subset}{\ensuremath{\subsetneq}}
\newcommand{\boldemph}{\emph}
\newcommand{\lcm}{\operatorname{lcm}}
\newcommand{\probskip}{\vskip1cm}
\newcommand{\subject}{MATH\xspace}
\newcommand{\coursenumber}{3140\xspace}
\newcommand{\semester}{Fall 2018\xspace}

\begin{document}
\thispagestyle{empty}

\noindent \textbf{\subject \coursenumber Homework 8}

\medskip

\noindent {\bf Exercises:} 1 and 2 (below); Judson Ch 9. 6, 19, 22, 27, 31, 37, and {\bf one} of 40, 41, 42\\
{\bf Recommended:} 21, 23 (hint: it's false), 28, 38--42, 45, 46, 53, 54\\
{\bf Due date:} Friday, 11/02

\bigskip

\noindent The first few exercises require some definitions from lecture, 
repeated here for your convenience.

\vskip5mm

\noindent Let $\bA = \<A, F^{\bA}\>$ and $\bB = \<B, F^{\bB}\>$ be two algebras of the
same \emph{similarity type}.  That is, to each operation
symbol $f \in F$ there corresponds an operation $f^{\bA}$ defined on $\bA$ and an
operation $f^{\bB}$ defined on $\bB$.
Thus, the set of operations defined on $\bA$ is the set 
$F^{\bA} = \{f^{\bA} : f\in F\}$; similarly 
$F^{\bB} = \{f^{\bB} : f\in F\}$.

\vskip3mm

\noindent For example, any two groups $\bG$ and $\bH$ have the same similarity type. 
To emphasize this, we could denote the operations of these groups using
the precise (albeit somewhat awkward) notation of the previous paragraph, as follows:
\[
\bG  = \<G, \circ^{\bG}, \invG, e^{\bG}\> \quad \text{ and } \quad
\bH  = \<H, \circ^{\bH}, \invH, e^{\bH}\>.
\] 
Here $\circ^{\bG}$, $\invG$, and $e^{\bG}$ represent the \emph{interpretation
in} $\bG$ of the binary, unary (inverse), and  nullary (identity)
operations that a group must possess (similarly for $\bH$).  

\vskip5mm

\noindent An \emph{algebra homomorphism} (or simply \emph{homomorphism}), denoted by
$\varphi: \bA \rightarrow \bB$, is a function $\varphi$ with domain $A$ and
codomain $B$ that satisfies the following conditions: for each $f \in F$, if $f$
is an $n$-ary operation symbol, and if $a_1, \dots, a_n \in A$, then
\[
\varphi(f^{\bA}(a_1, \dots, a_n)) = f^{\bB}(\varphi(a_1), \dots, \varphi(a_n)).
\]

\vskip3mm

\noindent For example, a \emph{group homomorphism} $\varphi: \bG \rightarrow \bH$  is a function 
$\varphi$ with domain $G$ and codomain $H$ that satisfies, $\forall x, y \in G$,
\begin{enumerate}
\item  $\varphi(x\circ^{\bG} y) = \varphi(x) \circ^{\bH} \varphi(y)$,
\item  $\varphi(\invG(x)) = \invH(\varphi(x))$,
\item  $\varphi(e^{\bG}) = e^{\bH}$.
\end{enumerate}

\vskip5mm

\noindent The textbook defines a group \emph{isomorphism} to be a group homomorphism that is both
one-to-one and onto.  This definition is fine for algebraic structures (like
groups).  
It does not work, however, for relational structures, like posets. (See Exercise 2 below).
A definition that works for both algebraic and relational structures is the following:
A homomorphism $\varphi : \bA \rightarrow \bB$ is an \emph{isomorphism} if there
exists a homomorphism $\psi: \bB \rightarrow \bA$ that composes with
$\varphi$ to give the identity, that is,
$\varphi \circ \psi = \operatorname{id}_B$
and $\psi \circ \varphi= \operatorname{id}_A$. (Here, $\operatorname{id}_X$
denotes the identity function on the set $X$: $\operatorname{id}_X(x) = x$.)

\newpage

\noindent {\bf Exercises}

\begin{enumerate}[{\bf 1.}]
%% 1 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\item When discussing two groups, like $\bG$ and $\bH$ above,
  our textbook uses more convenient notation, such as 
  $(G, \cdot)$ and $(H, \circ)$ (or, even more simply, $G$ and $H$).  The book
  defines a \emph{homomorphism} to be a function $\varphi: G\rightarrow H$ satisfying
  $\varphi(x\cdot y) = \varphi(x) \circ \varphi(y)$.
  Prove that this is equivalent to the definition given above by showing that 
  conditions (2) and (3) are unnecessary.\\
  ({\it Hint.} Assuming (1), derive (3), then derive (2).)

\vfill

% %% 2 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \item
% Define a lattice homomorphism.
% Then consider a lattice $\bL = \<L, \meet, \join\>$ and a poset $\bP = \<P, \preccurlyeq\>$.
% Is it possible to define a homomorphism $\varphi: \bL \rightarrow \bP$?  Explain.
% 
% \vfill

%% 3 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\item
A \emph{poset homomorphism} is an order preserving map.  That is, if
$\bP = \<P, \leqslant\>$ and 
$\bQ = \<Q, \preccurlyeq\>$ are two partially 
ordered sets, then a homomorphism 
$\varphi: \bP \rightarrow \bQ$  is a function satisfying, for all $x, y\in P$, 
if $x \leqslant y$ then $\varphi(x)\preccurlyeq \varphi(y)$.
Consider the two definitions of \emph{isomorphism} given in the last paragraph
on Page 1 above.
Using the two posets shown below, explain why the first of these definitions is 
not appropriate for posets.

\vskip5mm

\begin{center}
\begin{tikzpicture}[scale=0.75]
  \node (0) at (0,.5) [fill,circle,inner sep=1.2pt] {};
  \node (1) at (2,3) [fill,circle,inner sep=1.2pt] {};
  \node (2) at (-2,3) [fill,circle,inner sep=1.2pt] {};
  \draw (0) node [right] {$a$};
  \draw (1) node [right] {$b$};
  \draw (2) node [right] {$c$};
  \node (3) at (0.5,-1) {$\bP = \<\{a,b,c\},  \leqslant\>$};
  \draw (1) to (0) to (2);
\end{tikzpicture}
\hskip3cm
\begin{tikzpicture}[scale=0.75]
  \node (0) at (0,0) [fill,circle,inner sep=1.2pt] {};
  \node (1) at (0,2) [fill,circle,inner sep=1.2pt] {};
  \node (2) at (0,4) [fill,circle,inner sep=1.2pt] {};
  \draw (0) node [right] {$0$};
  \draw (1) node [right] {$1$};
  \draw (2) node [right] {$2$};
  \node (3) at (0.5,-1.5) {$\bQ = \<\{0,1,2\}, \preccurlyeq\>$};
  \draw (0) to (1) to (2);
\end{tikzpicture}
\end{center}

\vfill

%% %% 7 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \item[{\bf 9.7.}] 
%% Show that any cyclic group of order $n$ is isomorphic to ${\mathbb Z}_n$. 

\newpage
\noindent {\bf Exercises from Judson.}

\medskip
\item[{\bf 9.6}]
Show that the $n$th roots of unity are isomorphic to ${\mathbb Z}_n$. 
\vfill 

\item[{\bf 9.19}]
Prove that $S_3 \times {\mathbb Z}_2$ is isomorphic to $D_6$. Can you
make a conjecture about $D_{2n}$? Prove your conjecture. [\emph{Hint:}
Draw the picture.] 

\vfill

\item[{\bf 9.22}]
Let $G$ be a group of order 20. If $G$ has subgroups $H$ and $K$ of
orders 4 and 5 respectively such that $hk = kh$ for all $h \in H$ and
$k \in K$, prove that $G$ is the internal direct product of $H$ and $K$. 

\vfill

% \item[{\bf 9.23}]
% Prove or disprove the following assertion. Let $G$, $H$, and $K$ be
% groups. If $G \times K \cong H \times K$, then $G \cong H$. 
% 
% \vfill

\item[{\bf 9.27}]
Let $G \cong H$. Show that if $G$ is cyclic, then so is $H$.

\vfill

\item[{\bf 9.31}]
Let $\phi : G_1 \rightarrow G_2$ and  $\psi : G_2 \rightarrow G_3$  be
isomorphisms. Show that  $\phi^{-1}$ and $\psi \circ \phi$ are both
isomorphisms. Using these results, show that the isomorphism of groups
determines an equivalence relation on the class of all groups.

\vfill 

% \item[{\bf 9.35.}]
% Prove that $a + ib \mapsto a - ib$ is an automorphism of ${\mathbb C}^*$. 
% \item[{\bf 9.36.}]
% Prove that $A \mapsto B^{-1}AB$ is an automorphism of $SL_2({\mathbb R})$
% for all $B$ in $GL_2({\mathbb R})$. 
% \item[{\bf 9.37.}]

\item[{\bf 9.37.}] An {\bf Automorphism} of a group $G$ is an isomorphism of $G$
with itself. Denote the set of all automorphisms of $G$ by
$\aut(G)$
Prove that  $\aut(G)$ is a subgroup of
$S_G$, the group of permutations of $G$. 

\vfill

\noindent Solve {\bf ONE} of the following. (Take your pick.)

% \medskip
% 
% \item[{\bf 9.38.}]
% Find $\aut( {\mathbb Z}_6)$.
 
% \item[{\bf 9.39.}]
% Find $\aut( {\mathbb Z})$.

\bigskip %probskip

% \item[{\bf 9.39.}]
% Find $\aut( {\mathbb Z})$.
% 
% \bigskip %probskip

\item[{\bf 9.40.}]
Find two nonisomorphic groups $G$ and $H$ such that $\aut(G) \cong \aut(
H)$. 
 
\bigskip %probskip

\item[{\bf 9.41.}]
Let $G$ be a group and $g \in G$. Define a map $i_g : G \rightarrow
G$\label{noteinner} 
by $i_g(x) = g x g^{-1}$.  Prove that $i_g$ defines an automorphism of
$G$.  Such an automorphism is called an \boldemph{inner
automorphism}\index{Automorphism!inner}. The set of all inner
automorphisms is denoted by $\inn(G)$\label{noteinneraut}. 

\bigskip %probskip
 

\item[{\bf 9.42.}]
Prove that $\inn(G)$ is a subgroup of $\aut(G)$.
 
 
\end{enumerate}
\end{document}
